---
title: "Employee Turnover Report"
date: "`r Sys.Date()`"
author: "Taren Shaw"
output:
  pdf_document:
    latex_engine: xelatex
sansfont: Calibri Light
---
```{r include=FALSE}
# Load Packages
pacman::p_load(tidyverse, readr, lubridate, skimr)
```

```{r include=FALSE}
# Load Data
df <- readr::read_csv("HRDataset_v14.csv")
```

```{r include=FALSE}
# Transform Data
df <- df %>% mutate(
  DateofTermination = mdy(DateofTermination),
  DateofHire = mdy(DateofHire))
```


```{r include=FALSE}
# Find min/max termination date
min_max <- df %>%
  drop_na(DateofTermination) %>%
  summarise(min = min(DateofTermination), max = max(DateofTermination)) 

min_year <- min_max$min %>% year() %>% as.character()
max_year <- min_max$max %>% year() %>% as.character()

min_period <- min_max$min %>% as.character()
max_period <- min_max$max %>% as.character()

```


# Introduction

This report serves to summarise the situation of employee turnover in the company for the years `r min_year` to `r max_year`. It is recommended that the HR council takes into account the conclusions garnered from the statistical anlysies generated from the study of the company's headcount reports which were downloaded from the HR system in use. This report serves to bring to the attention the current and past situation of employee turnover within the company to the HR council and hopes to help with creating a plan of action if it should be necessary.

1. This report has been prepared to provide an up to date analysis of employee turnover and as part of the Business's workforce planning considerations.

2. Employee turnover has been assessed on the basis of the number of employees leaving the Company as a percentage of the total number of staff (headcount) employed by the Company.

3. This report presents the turnover figures between `r min_period` (date of the first termination) and `r max_period` (date of the last termination).

# Analysis

## Emplyee Turnover Rate

Before presenting the analysis, it is important to define what employee turnover is and show the method for calculating yearly employee turnover. Turnover rate is the percentage of employees in a workforce that leave during a certain period of time, which includes involuntary and voluntary terminations (as defined in the HR system in use). In this report we have calculated it using the following formula:

$$\ AnnualTurnoverRate = \frac{LeaversInPeriod}{(BeginningEmployeesInPeriod + EndingEmployeesInPeriod)/2}*100   $$


### Leavers

We have defiend leavers for this study as all terminations found within the HR system that are involuntray and voluntray. These include terminations such as termination, retirement, death, interagency transfers, and resignations. 

In figure x we have aggregated all terminations by year, starting from the first recorded termination in the system on `r min_period` to the most recent termination recorded on `r max_period`.

```{r include=FALSE}
# Define years range
years <- seq(from = year(min_max$min), to = year(min_max$max), by = 1) %>%
  as.character()
```

```{r include=FALSE}
n_leavers <- function(year = "2010") {
  df %>%
    filter(year(DateofTermination) == year) %>%
    nrow()
}
```

```{r echo=FALSE}
var_turn_plt <- function(x, title = "Title") {

  # find the median of x
  
  median <- map_dbl(years, x) %>%
    setNames(years) %>%
    as.data.frame() %>%
    setNames("n") %>%
    rownames_to_column("Year") %>%
    summarise(median = median(n)) %>%
    as.numeric()
  
  # Plot df with median
  plt <- map_dbl(years, x) %>%
    setNames(years) %>%
    as.data.frame() %>%
    setNames("n") %>%
    rownames_to_column("Year") %>%
    ggplot(aes(x = .[, 1], y = .[, 2], group = 1)) +
    geom_line(color = "#0072b1") +
    geom_point(color = "#0072b1") +
    geom_hline(yintercept = median, linetype = "dashed", color = "red") +
    theme_bw() +
    geom_text(aes(label = n), vjust = -0.5) +
    labs(title = title) +
    xlab("Year") +
    ylab("n")

}

n_leavers_plt <- var_turn_plt(n_leavers, "Terminations")
n_leavers_plt
```

```{r include=FALSE}
# Dataframes of Variabel functions
df_count <- function(x) {
  map_dbl(years, x) %>%
    setNames(years) %>%
    as.data.frame() %>%
    setNames("n") %>%
    rownames_to_column("Year") %>% 
    tibble()
}
```


```{r include=FALSE}
median <- map_dbl(years, n_leavers) %>%
    setNames(years) %>%
    as.data.frame() %>%
    setNames("n") %>%
    rownames_to_column("Year") %>%
    summarise(median = median(n)) %>%
    as.numeric()

```

```{r include=FALSE}
max_n_leaves_year <- df_count(n_leavers) %>% filter(n == max(n)) %>% select(Year) %>% as.character()
max_n_leavesr <- df_count(n_leavers) %>% filter(n == max(n)) %>% select(n) %>% as.character()

Eval1 <- if (max_n_leaves_year == max_year) {
"one"
} else {
"two"
}  

y <- paste("As shown in fig x we can see that the current year ", max_n_leaves_year, " has the highest number of leavers"," (", max_n_leavesr ,") ", "compared to the previous years. As the number of leavers is greater than previous years, this report recommends that the HR council investigates this matter further, not only using a descriptive analysis, but also a diagnostic one. We recommend that a deeper study is initiated into where these leavers are coming from in the company, what type of leavers they are (voluntary or involuntary), and other compounding factors.", sep = "")

x <- "Test"
```


`r if(Eval1 == "one"){y}`
`r if(Eval1 == "two"){x}`




